Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion

### 2023-08-22 Manuscript.Rmd
updating figures

source(here::here("scripts","utility.R"))
source(here::here("scripts","plots.R"))
source(here::here("scripts","supplementary.R"))

#get models
source(here::here("scripts","models.R"))

@@

list2env(stats, envir=environment())

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 1: A and B) Placement and distance of thermodes on the inner forearm, showing corresponding spinal mapping for all four conditions. Within dermatomes, the relative location of the cold thermode was proximal or distal. Across dermatomes, the relative location of the corresponding spinal segments was rostral or caudal. Within dermatome conditions also included warm and cold thermodes in C6 (not depicted here), as well as T1. C) The three VAS participants used to report sensation coming from the reference probe, presented for each trial in a randomised order."}
ggdraw() +
    draw_image(magick::image_read(here::here("figures","figure1_v2.PNG")), scale = 1)
```

@@

The choices of the zero-inflated approach and the use of beta regressions were necessitated by the specific distribution of VAS ratings.
The beta distribution is suitable for modelling VAS rating data, as they are proportional in nature.
Additionally, the zero-inflation was needed due to the presence of an excess number of zero values in specific ratings and conditions.
Specifically, we anticipated an over-representation of zero values for thermosensory ratings that were counterfactual to the objective stimulation quality (i.e., cold ratings of warm stimuli and warm ratings of cold stimuli) and burning ratings of non-TGI stimuli.
The latter stimuli were designed to not elicit an illusion or trigger a weaker illusion as compared to the TGI stimuli.
We carried out the statistical analyses using the \'glmmTMB\' package in R (version `r sessionInfo()$otherPkgs$glmmTMB$Version`), and statistical significance was set at p \< .05.

@@

We would like to thank XXX for their help with participant recruitment and data collection.
This study was supported by a European Research Council Starting Grant (ERC-2020-StG-948788).

@@

```{r,warning = F, message = F, echo = F}
get_main_tables(model_cold_exp1,round = 2, title = "Exp. 1: Cold reference on cold ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")

get_main_tables(model_cold_exp2,round = 2, title = "Exp. 2: Warm reference on cold ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@

```{r,warning = F, message = F, echo = F}

get_main_tables(model_warm_exp1,round = 2, title = "Exp.1: Cold reference on warm ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")

get_main_tables(model_warm_exp2,round = 2, title = "Exp. 2: Warm reference on warm ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@

```{r,warning = F, message = F, echo = F}

get_main_tables(model_burn_exp1,round = 2, title = "Exp. 1: Cold reference on burning ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")

get_main_tables(model_burn_exp2,round = 2, title = "Exp. 2: Warm reference on burning ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all") %>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@

Experiment 1: warm ratings (stimulation by dermatome interaction: \$\\beta\$ = r stats_model_warm_exp1\$beta[8], r make_pvalue(stats_model_warm_exp1\$p[8]); dermatome main effect: $\beta$ = `r stats_model_warm_exp1$beta[4]`, `r make_pvalue(stats_model_warm_exp1$p[4])`) or burning ratings (stimulation by dermatome interaction: $\beta$ = `r stats_model_burn_exp1$beta[8]`, `r make_pvalue(stats_model_burn_exp1$p[8])`; dermatome main effect: $\beta$ = `r stats_model_burn_exp1$beta[4]`, `r make_pvalue(stats_model_burn_exp1$p[4])`)

Experiment 2: warm ratings (stimulation by dermatome interaction: $\beta$ = `r stats_model_warm_exp2$beta[8]`, `r make_pvalue(stats_model_warm_exp2$p[8])`); dermatome main effect: $\beta$ = `r stats_model_warm_exp2$beta[4]`, `r make_pvalue(stats_model_warm_exp2$p[4])`) or burning ratings (stimulation by dermatome interaction: $\beta$ = `r stats_model_burn_exp2$beta[8]`, `r make_pvalue(stats_model_burn_exp2$p[8])`); dermatome main effect: $\beta$ = `r stats_model_burn_exp2$beta[4]`, `r make_pvalue(stats_model_burn_exp2$p[4])`).

Experiment 1: burn ratings (stimulation by rostrocaudal location: $\beta$ = `r stats_model_burn_exp1$beta[7]`, `r make_pvalue(stats_model_burn_exp1$p[7])`; rostrocaudal main effect ($\beta$ = `r stats_model_burn_exp1$beta[3]`, `r make_pvalue(stats_model_burn_exp1$p[3])`)

>>>>

# functions and data
source(file.path("scripts","utility.R"))
source(file.path("scripts","plots.R"))
source(file.path("scripts","supplementary.R"))
source(file.path("scripts","models.R"))

@@

list2env(stats, envir=environment())
```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 1: A and B) Placement and distance of thermodes on the inner forearm, showing corresponding spinal mapping for all four conditions. Within dermatomes, the relative location of the cold thermode was proximal or distal. Across dermatomes, the relative location of the corresponding spinal segments was rostral or caudal. Within dermatome conditions also included warm and cold thermodes in C6 (not depicted here), as well as T1. C) The three VAS participants used to report sensation coming from the reference probe, presented for each trial in a randomised order."}
ggdraw() +
    draw_image(magick::image_read(file.path("figures","figure1_v2.PNG")), scale = 1)
```

@@

The choices of the zero-inflated approach and the use of beta regressions were necessitated by the specific distribution of VAS ratings.
The beta distribution is suitable for modelling VAS rating data, as they are proportional in nature.
Additionally, the zero-inflation was needed due to the presence of an excess number of zero values in specific ratings and conditions.
Specifically, we anticipated an over-representation of zero values for thermosensory ratings that were counterfactual to the objective stimulation quality (i.e., cold ratings of warm stimuli and warm ratings of cold stimuli) and burning ratings of non-TGI stimuli.
The latter stimuli were designed to not elicit an illusion or trigger a weaker illusion as compared to the TGI stimuli.
We carried out the statistical analyses using the 'glmmTMB' package in R (version `r sessionInfo()$otherPkgs$glmmTMB$Version`), and statistical significance was set at p \< .05.

@@

We thank Małgorzata Basińska, Patrik Molnár, Sára A. Szabó for their help with participant recruitment and data collection.
We also thank Michael Clements for the collection of pilot data that was instrumental in refining the experimental design.
This study was supported by a European Research Council Starting Grant (ERC-2020-StG-948788).

@@

Tables of main effects and interactions for each sensory quality (cold, warm, burning) for experiment 1 and experiment 2.

## Cold ratings

```{r,warning = F, message = F, echo = F}
get_main_tables(model_cold_exp1,round = 2, title = "Exp. 1: Cold reference on cold ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

```{r,warning = F, message = F, echo = F}
get_main_tables(model_cold_exp2,round = 2, title = "Exp. 2: Warm reference on cold ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@

## Warm ratings

```{r,warning = F, message = F, echo = F}
get_main_tables(model_warm_exp1,round = 2, title = "Exp.1: Cold reference on warm ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

```{r,warning = F, message = F, echo = F}
get_main_tables(model_warm_exp2,round = 2, title = "Exp. 2: Warm reference on warm ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@

## Burning ratings

```{r,warning = F, message = F, echo = F}
get_main_tables(model_burn_exp1,round = 2, title = "Exp. 1: Cold reference on burning ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all")%>% flextable::font(fontname = "Times New Roman", part = "all")
```

```{r,warning = F, message = F, echo = F}
get_main_tables(model_burn_exp2,round = 2, title = "Exp. 2: Warm reference on burning ratings") %>%
  width(j = 1, width = 2.5) %>%
  width(j = c(2,3,4,5), width = 1.1)%>%
  fontsize(size = 10, part = "all") %>% flextable::font(fontname = "Times New Roman", part = "all")
```

@@
